---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
activity <- read.csv(unzip("activity.zip"))
summary(activity)
str(activity)
```

Let's obtain clean step data:

```{r}
cl_steps <- activity$steps[!is.na(activity$steps)]
```

## What is mean total number of steps taken per day?

1. A histogram of the total number of steps taken each day:

```{r}
library(ggplot2)
qplot(cl_steps, binwidth = diff(range(cl_steps))/100)
qplot(cl_steps, binwidth = diff(range(cl_steps))/100)+coord_cartesian(ylim=c(0, 500))
```

And one without zeros:
```{r}
qplot(cl_steps, binwidth = diff(range(cl_steps))/100, xlim=c(1, 800))
```


```{r}
mean(activity$steps, na.rm = TRUE)
median(activity$steps, na.rm = TRUE)
```


## What is the average daily activity pattern?

```{r}
avg <- function(intrv, df) {
  mean(df$steps[df$interval==intrv], na.rm = TRUE)
}

ints <- unique(activity$interval)
stats <- list(intervals=ints, avg_steps=sapply(ints, avg, activity))

plot(stats$intervals, stats$avg_steps, type="l")
```

```{r}
max_steps <- max(stats$avg_steps)
max_int <- stats$intervals[stats$avg_steps == max_steps]
```
Max count of steps (`r max_steps`) is achieved on interval #`r max_int`.


## Imputing missing values

Number of NAs:
```{r}
sum(is.na(activity$steps))
```

Replacing NAs with intervals averages:
```{r}
restored <- activity
sum(is.na(restored$steps))

nas <- is.na(activity$steps)

for(i in 1:length(nas)){
  if(nas[i]){
    restored$steps[i] <- stats$avg_steps[stats$interval == restored$interval[i]]
  }
}

# restored$steps[nas] <- stats$avg_steps[which(stats$interval %in% restored$interval[nas])]

sum(is.na(restored$steps))

qplot(steps, data = restored, binwidth = diff(range(cl_steps))/100)

```


```{r}
mean(restored$steps)
median(restored$steps)
```



## Are there differences in activity patterns between weekdays and weekends?


```{r}
is_weekend <- function(date){
  w <- weekdays(date)
  w =="Saturday" || w == "Sunday"
}

restored$wtype <- sapply(restored$date, function(d){
  factor(levels = c("weekday", "weekend"),
            x = if(is_weekend(as.Date(d))){"weekend"} else {'weekday'})
  })

sum(restored$wtype == "weekend")

sum(restored$wtype == "weekday")

weekends <- restored[restored$wtype == "weekend",]
attr(weekends, 'name') <- "Weekends"
weekdays <- restored[restored$wtype == "weekday",]
attr(weekdays, 'name') <- "Weekdays"

par(mfrow = c(2, 1))

for(df in list(weekdays, weekends)){
  str(df)
  ints <- unique(df$interval)
  stats <- list(intervals=ints, avg_steps=sapply(ints, avg, df))
  print(any(is.na(stats)))
  plot(stats$intervals, stats$avg_steps, type="l", main=attributes(df)$name)
}

```